---
title: "Reproducible Research: Peer Assessment 1"
author: "JBF"
date: "7/1/2016"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r}
download.file(url="http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",
              destfile="repdata data activity.zip")
unzip(zipfile="repdata data activity.zip")

data <- read.csv(file="activity.csv")
data$date <- as.Date(data$date, format="%Y-%m-%d")
data$hour <- as.factor(data$interval)
levels(data$hour) <- sprintf("%2d:%s", rep(0:23,each=12), 
                             rep(c("00","05",seq(10,55,5)),24))
data$time <- strptime(paste(data$date,data$hour), format="%Y-%m-%d %H:%M")
```

## What is mean total number of steps taken per day?
```{r}
library(dplyr)
data_day <- data %>% select(steps, date) %>% group_by(date) %>%
  summarise(total=sum(steps,na.rm=T))
data_day_summ <- data_day %>% summarise(mn=mean(total), md=median(total))
```

The histogram of the total number of steps taken per day
```{r}
library(ggplot2)
ggplot(data_day, aes(x=total)) + geom_histogram()
```

The mean of total number of steps taken per day is `r round(data_day_summ$mn,2)` and the median is `r data_day_summ$md`.

## What is the average daily activity pattern?
```{r}
data_interval <- data %>% select(steps, interval, hour) %>% 
  group_by(interval, hour) %>% summarise(mn=mean(steps, na.rm=T))
ggplot(data_interval, aes(x=interval, y=mn)) + geom_line() + 
  scale_x_continuous(breaks=data_interval$interval[seq(1,288,24)],
                     labels=data_interval$hour[seq(1,288,24)])
int_max <- as.character(data_interval$hour[which.max(data_interval$mn)])
```

The interval `r int_max` contains the maximum number of steps.

## Imputing missing values
```{r}
num_mis <- sum(is.na(data$steps))
```
The total number of `r num_mis`.

We impute with the mean of each interval across all days.
```{r}
data_imp <- data %>% select(-time) %>% left_join(data_interval)
data_imp$steps[is.na(data_imp$steps)] <- data_imp$mn[is.na(data_imp$steps)]
```

The total number of steps per day for the imputed dataset.
```{r}
data_day_imp <- data_imp %>% select(steps, date) %>% group_by(date) %>%
  summarise(total=sum(steps,na.rm=T))
data_day_imp_summ <- data_day_imp %>% summarise(mn=mean(total), md=median(total))
```

The histogram of the total number of steps taken per day for the imputed dataset.
```{r}
ggplot(data_day_imp, aes(x=total)) + geom_histogram()
```

The mean of total number of steps taken per day is `r round(data_day_imp_summ$mn,2)` and the median is `r data_day_imp_summ$md`.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
data_imp$type_day <- ifelse(weekdays(data_imp$date) %in% c("Saturday","Sunday"),
                            "weekend", "weekday")
data_imp$type_day <- as.factor(data_imp$type_day)
data_imp_interval <- data_imp %>% select(steps, interval, hour, type_day) %>% 
  group_by(type_day, interval, hour) %>% summarise(mn=mean(steps, na.rm=T))
ggplot(data_imp_interval, aes(x=interval, y=mn)) + geom_line() + 
  scale_x_continuous(breaks=data_interval$interval[seq(1,288,24)],
                     labels=data_interval$hour[seq(1,288,24)]) +
  facet_grid(type_day ~ .)
```

